---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# gendercodeR

The goal of gendercodeR is to allow simple recoding of freetext gender responses.

## Why would we do this?

Researchers who collect self-reported demographic data from respondents occasionally collect gender using a free-text response option. This has the advantage of respecting the gender diversity of respondents without prompting users and potentially including misleading responses. However, this presents a challenge to researchers in that some inconsistencies in typography and spelling create a larger set of responses than would be required to fully capture the demographic characteristics of the sample. 

For example, male participants may provide freetext responses as "male", "man", "mail", "mael". Non-binary participants may provide responses as "nonbinary", "enby", "non-binary", "non binary"

This package uses dictionaries of common misspellings to recode these freetext responses into a consistent set of responses.

## Installation

This package is not on CRAN. To use this package please run the following code:


``` r
devtools::install_github("ropenscilabs/gendercodeR")
library(gendercodeR)
```

## Example

You have a dataframe with the following format where individuals have provided 
gender in a range of inconsistent formats. 

```{r example}

library(gendercodeR)

df <- data.frame(stringsAsFactors=FALSE,
      gender = c("male", "MALE", "mle", "I am male", "femail", "female", "enby"),
         age = c(34L, 37L, 77L, 52L, 68L, 67L, 83L)
)

df
```

Running the `genderRecode()` function will take the inputted dataset, match 
freetext gender responses to the dictionary, create a new column in the 
dataframe with the recoded gender response. For freetext gender responses that
are not in the dictionary the original freetext will be copied to the recoded 
gender column.

```{r} 
genderRecoded <- genderRecode(input=df,
                              genderColName = "gender", 
                              method = "broad",
                              outputColName = "gender_recode", 
                              missingValuesObjectName = NA,
           customDictionary = NULL)

genderRecoded
```
## Options within the function

### method

The package provides the option to either correct spelling and standardise terms 
while maintaining the diversity of responses. This is selected by setting 
`method =  "broad"` or to compress all responses down to male/female/'sex and 
gender diverse' using `method =  "narrow"`.

##### Example using narrow coding 
```{r} 
genderRecoded <- genderRecode(input=df,
                              genderColName = "gender", 
                              method = "narrow",
                              outputColName = "gender2", 
                              missingValuesObjectName = NA,
           customDictionary = NULL)

genderRecoded
```

### missingValuesObjectName

By default the unmatched responses and the number of times it exists in the 
input dataframe. Setting `missingValuesObjectName` saves the list of unmatched 
free text gender responses. We recommend assessing these as they may be 
human code-able (e.g. "I am male") or otherwise meaningful and not captured by 
our base dictionary. We expect that researchers may wish to manually recode 
such responses or add frequently occurring responses to their custom dictionary.

### customDictionary

Users can specify a custom dictionary that supplements or overwrites entries
to the in-built dictionary. 

There is a vignette for this function

## Viewing the inbuilt dictionary

The inbuilt dictionary can be assigned to an object using the `getDictionary()` 
function. This function does not take any arguments.

```{r} 
test <- getDictionary()
test
```


# Contributing to this package

This package is a reflection of cultural context of the package contributors we 
welcome issues and pull requests to make the package more inclusive and/or 
suitable for a broader range of cultural contexts. 

# Acknowledgement of Country

We acknowledge the Wurundjeri people of the Kulin Nation as the custodians of 
the land on which this package was developed and pay respects to elders past, 
present and future.
