---
title: "Learning R Survey Results"
author: "Carl Howe, RStudio"
date: "14 December 2018"
output:
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
    logo: rstudio.png
    smaller: yes
    widescreen: yes
subtitle: 14 December 2018
params:
  bar_colors: dark green
  language: English
---

# Learning R Survey Results

## Methodology

- Survey conducted between December 6 and December 31, 2018
- Survey was fielded in both English and Spanish versions
- Respondents solicited from
    + community.rstudio.com
    + Twitter followers of RStudio employees
    + reddit.com/datascience
- Survey results are not representative of any broader population
- Data and processing scripts can be found the repo at the end of this talk
- Data and scripts are open source

## Setup

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(googlesheets)
library(choroplethr)
library(choroplethrMaps)
data(country.map)
library(tidyverse)
library(RColorBrewer)
library(gendercodeR)
library(ggrepel)

###
### Let's define helper functions before we get going:
###

# This tallies up the results for a question
# rm.na = TRUE to calculate percentages based on non-NA results; set to FALSE to include NAs
# set sort = TRUE sorts the result by count; set to FALSE to return in whatever order they occur

tally_question <- function(df, question_name, rm.na = TRUE, sort = TRUE) {
  quoted_question <- enquo(question_name)
  filtered_df <- df
  if (rm.na) {
    filtered_df <- filtered_df %>% filter(!is.na(!!quoted_question))
  }
  results_df <- filtered_df %>% 
    count(!!quoted_question, sort = sort) %>% add_tally() %>% 
    mutate(percent = round(n / nn * 100))
  return(results_df)
}


# Split and aggregate: derives multiple answers to a single question by separating on commas and returning
# the results as an embedded list in the dataframe.
split_and_aggregate <- function(df, question_name) {
  quoted_question <- enquo(question_name)
  responses_df <- df %>% summarize(responses = sum(!is.na(!!quoted_question)))
  splits <- df %>% mutate(items = purrr::map(!!quoted_question, str_split, ", ")) %>% unnest()
  
  aggregated_items <- splits %>% unnest() %>% group_by(items) %>% count(sort=TRUE)
  aggregated_items <- aggregated_items %>% mutate(num_responses = responses_df$responses)
  return(aggregated_items)
}

# Top N choices: function to distill many possible results to a question to the top N
# responses, with the rest aggregated into an "Other" answer.
top_n_choices <- function(df, column_name, total_responses, num = 10) {
  
  quoted_column_name <- enquo(column_name)
  summarized_responses <- df %>% 
    mutate(percent = round(n / total_responses * 100)) %>% 
    arrange(desc(percent))

  # Now take these responses and only show the top N, aggregating the rest into an Other category

  literals <- head(summarized_responses, num) %>% 
    ungroup()
  other <- tail(summarized_responses, -num) %>% 
    ungroup() %>% 
    summarize(!!quoted_column_name := "Other", 
              n = sum(n),
              percent = round(n / first(total_responses) * 100))
  top_n <- rbind(literals, other) %>% drop_na()
  return(top_n)
}

question_text <- function(question_name_string, wrap_length = 55)
{
   question_text <- survey_questions %>% filter(Question_name == question_name_string) %>% select(Question_text)
   if (str_length(question_text$Question_text) >= wrap_length) {
     return_text <- question_text$Question_text %>% str_wrap(width = wrap_length - 5)
   } else {
     return_text <- question_text$Question_text
   }
   return(paste0('"', return_text, '"'))
}

## Plot a pie chart from a single response question. The call form is a little baroque, so here's an example
## to illustrate the mapping of values to arguments.
##     pie_chart(df = ethnicities, 
##               column_name            = Qethnicity_coded, 
##               original_column_string = "Qethnicity",
##               text_position          = position        ## the name of the position column in the df
##               title_string           = "Ethnicities", 
##               fill_legend_title      = "Respondent Ethnicity"
##               colors                 = "Set1"         From the RColorBrewer palettes

pie_chart <- function(df, 
                      column_name, 
                      original_column_string, 
                      text_position,
                      title_string, 
                      fill_legend_title, 
                      colors ) {
  quoted_column_name <- enquo(column_name)
  quoted_position <- enquo(text_position)
  p <- ggplot(df) +
    geom_col(aes(x = 1, y = percent, fill = !!quoted_column_name), 
             width = 1, size = 1, color = "gray90", alpha = 0.4) +
    geom_text(aes(x = 1.2, y = !!quoted_position, label = label),
                    size = 3,
                    hjust = 0.4,
                    color = "black"
     ) +
    labs(x = NULL, y = NULL, 
         title = title_string,
         subtitle = question_text(original_column_string),
         caption = "2018 RStudio Learning R Survey",
         fill = fill_legend_title) +
    guides(fill = guide_legend()) +
    scale_fill_brewer(palette = colors) +
    #   scale_fill_viridis_d("Ethnicities", option = "A", direction = -1, alpha = 0.4) +
    theme(legend.key.width=unit(1, "cm")) +
    theme_classic() +
    coord_polar("y") +
    #    coord_flip() +
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
  print(p)
}

```


```{r read in survey, include = FALSE}
##############################################################
### This section begins the mainline code to read the survey #
##############################################################

# First, we'll just read in the data and fix all the question names, which in Google Forms, are literally
# the full text of the question. I've put short question names on another sheet of the results Google Sheet, so
# we do the substitution here.

mysheets <- gs_ls() # will authenticate

sheet_name <- case_when(
  params$language == "English" ~ "Learning R Internet Survey",
  params$language == "Spanish" ~ "Learning R Internet Survey (Spanish Version)",
  TRUE ~ "Unknown language"
)

if (sheet_name == "Unknown language") {
  print ("YAML header parameter not set to Spanish or English")
  stop();
}

survey_handle <- gs_title(sheet_name)

# get the survey results
survey <- survey_handle %>%
  gs_read(ws = "Form Responses 1")

## Unfortunately, this read brings in the question text as all the column names. That's not good.
## We'll save them in question_text and put shorter, more friendly names in instead.

survey_questions <- survey_handle %>% gs_read(ws="Question Names")
names(survey) <- survey_questions$Question_name
respondents <- nrow(survey)
survey_save <- survey                # in case I mess up survey, I can check it against this
```


```{r survey pre-processing, include=FALSE}

survey <- survey_save   # This is just to ensure I get the original survey each time I run this block.

# Make the times correct times
est <- locale(tz="US/Eastern")
survey$Qtime <- parse_datetime(survey$Qtime, format = "%m/%d/%Y %H:%M:%S", locale=est)

# This learner_type variable defines if someone is a recent R learner or not for faceting
survey <- survey %>% 
  mutate(learner_type = ifelse(Qr_year < 1900, NA, ifelse(Qr_year <= 2016, "Early Learner", "Recent Learner")))

# This codes the open-text gender question into a more friendly form for analysis
survey$Qgender <- survey$Qgender %>% tolower() %>% str_trim() %>% str_replace_all("[[:punct:]]", "")

opentext_gender_dictionary <- read_csv("data/opentext_gender_dictionary.csv")
gender_dictionary <- opentext_gender_dictionary %>%  mutate(Input = str_replace_all(Input, "[[:punct:]]", ""))
survey <- genderRecode(survey, method = "narrow", 
                       genderColName = "Qgender",
                       outputColName = "Qgender_coded",
                       customDictionary = gender_dictionary)
survey <- survey %>% mutate(Qgender_coded = ifelse(Qgender_coded == "", NA, Qgender_coded))
uncoded_genders <- survey %>% group_by(Qgender_coded) %>% count(sort = TRUE)
# write_csv(uncoded_genders, "data/uncoded_genders_language.csv")  # For detecting open ends we should code

# This codes the ethnicity question for further analysis
# We recognize that this is both an offensive question to some and an imperfect classification
# Its purpose is to simply measure how diverse our community is. Our rationale is that if
# we do not attempt to measure the R community's diversity, then our claims of community diversity
# are meaningless because our only measurements are anecdotal.

opentext_ethnicity_dictionary <- read_csv("data/opentext_ethnicity_dictionary.csv")
ethnicity_dictionary <- opentext_ethnicity_dictionary %>%  
  mutate(Input = str_replace_all(Input, "[[:punct:]]", ""))

# Yes, I know this will cause false categorizations of open-ended responses with commas as Multiple Ethnicities
# That said, I can't see an obvious other way to detect multiple ethnicities beyond the dictionary matches below

survey <- survey %>% 
  mutate(Qethnicity_processed = ifelse(str_detect(Qethnicity, ","), "Multiple Ethnicities", Qethnicity))
survey$Qethnicity_processed <- survey$Qethnicity_processed %>% tolower() %>% str_trim() %>% str_replace_all("[[:punct:]]", "")
survey <- survey %>% 
  left_join(ethnicity_dictionary, by=c("Qethnicity_processed" = "Input"))
uncoded_ethnicities <-survey %>% 
  anti_join(ethnicity_dictionary, by=c("Qethnicity_processed" = "Input")) %>% 
  count(Qethnicity_processed, sort = TRUE) 
survey <- survey %>% 
  mutate(Qethnicity_coded = ifelse(Qethnicity_coded == "Prefer not to answer",
                                   NA, 
                                   Qethnicity_coded))
collected_ethnicities <- survey %>% group_by(Qethnicity_coded) %>% count(sort = TRUE)
theme_set(theme_classic())

```

## Responses over time

```{r responses over time}
survey <- survey %>% mutate(number_responses = cumsum(!is.na(Qtime)))
ggplot(survey) + 
  geom_line(aes(x=Qtime, y=number_responses)) + 
  ggtitle(sprintf("%s Survey Responses Over Time (N = %d)",params$language, respondents)) + 
  labs(y = "Time", x = "Response Date")

```

# Demographics

## Respondent Ages

```{r Respondent Ages, echo = TRUE}

ggplot(survey) +
  geom_histogram(aes(2018 - Qyear_born), binwidth = 5,  fill = params$bar_colors, color="black") +
  scale_x_continuous(breaks = seq(15,80,5)) +
  labs(title = "Respondent Ages (Histogram)", x="Age", y="Count")
```

## Respondent Gender

```{r Respondent Gender, echo = TRUE}

genders <- survey %>% 
  group_by(Qgender_coded) %>% 
  count(sort = TRUE) %>% 
  mutate(unit = 1) %>% 
  drop_na()

gender_responses <- sum(genders$n, na.rm = TRUE)
genders <- genders %>% 
  group_by(unit) %>% 
         mutate(percent = round(n / gender_responses * 100),
         position = cumsum(percent) - percent / 2,
         label = ifelse(percent >= 10, paste0(Qgender_coded, "\n", round(percent), "%"), "")) %>% 
  arrange(desc(percent))
genders <- genders %>% 
  mutate(Qgender_coded = factor(as.character(Qgender_coded), levels = c("female", "male", "sex and gender diverse", "unclear")))
    
pie_chart(df                 = genders,
          column_name        = Qgender_coded,
          original_column_string = "Qgender",
          text_position          = position,
          title_string           = "Respondent Gender",
          fill_legend_title      = "Gender",
          colors                 = "Set1")

# ggplot(genders, aes(x = "", y = percent, fill = Qgender_coded)) +
#     geom_bar(width = 1, size = 1, color = "gray90", stat = "identity", alpha = 0.7) +
#     coord_polar("y") +
#     geom_text(aes(label = paste0(round(percent), "%")), 
#               position = position_stack(vjust = 0.5)) +
#     labs(x = NULL, y = NULL, fill = NULL, 
#          title = "Respondent Gender", 
#          subtitle = question_text("Qgender"), 
#          caption = "2018 RStudio Learning R Survey") +
#     guides(fill = guide_legend()) +
#     scale_fill_brewer(palette = "Set1") +
#     theme_classic() +
#     theme(axis.line = element_blank(),
#           axis.text = element_blank(),
#           axis.ticks = element_blank())

```

## Respondent Countries

```{r Respondent Countries, echo = TRUE}
countries <- survey %>% count(Qcountry, sort=TRUE) %>% drop_na()
names(countries) <- c("region", "value")
countries <- countries %>% mutate(region = tolower(region))
unknown_countries <- c("singapore", "korea", "south", "serbia and montenegro", "hong kong", "andorra", "bahrain", "christmas island", "macau", "mauritius", "puerto rico", "tanzania")
choro_countries <- countries %>% filter(!(region %in% unknown_countries))
choro_countries <- choro_countries %>% mutate(region = str_replace_all(region, "korea, south", "south korea"))

# country_choropleth(choro_countries, title = "", legend = "", num_colors = 8,
#  zoom = NULL)

plot_countries <- country.map %>% left_join(choro_countries, by = "region")
ggplot(data = plot_countries, aes(long, lat, group=group)) + 
  geom_polygon(aes(fill = value), color = "gray", size=0.1) +
  scale_fill_gradient(trans = "log10", low = "dark blue", high = "yellow") +
  coord_fixed(1.3) +
  scale_x_continuous(breaks = seq(-150, 150, 60)) +
  scale_y_continuous(breaks = seq(-60, 90, 30), limits = c(-60, 90)) +
  labs(title = "Geographical Distribution of Survey Respondents", 
       subtitle = question_text("Qcountry"),
       caption = "2018 RStudio Learning R Survey",
       x = "", y = "", fill = "Responses")
```

## Industry

```{r Respondent industry, echo = TRUE}
industry <- survey %>% count(Qindustry, sort = TRUE)
industry_responses <- sum(!is.na(survey$Qindustry))
industry <- industry %>% mutate(percent = round(n / industry_responses * 100, 0))

top_10_industries <- top_n_choices(industry, Qindustry, industry_responses, 10)

# ggplot(industry, aes(area = percent, fill = as.numeric(percent), label = sprintf("%s\n%d%%", Qindustry, percent))) +
#  geom_treemap() +
#  geom_treemap_text(color = "white", 
#                    place = "center",
#                    reflow = TRUE) +
#  scale_fill_continuous(type = "gradient") +
#  labs(title = "Respondent Industry")

top_10_industries <- top_10_industries %>% 
  mutate(short_industry = str_replace(Qindustry, " \\(Federal, State, or Local\\)", ""),
         short_industry = str_replace(short_industry, "Services", "Svcs."),
         short_industry = str_replace(short_industry, "Information Technologies", "Info. Tech."),
         short_industry = str_replace(short_industry, "Professional and ", "Prof./"),
         short_industry = str_replace(short_industry, " and Activities", ""),
         short_industry = str_replace(short_industry, " and Medicine", ""),
         short_industry = str_replace(short_industry, " Resources and ", " Resources/"),
         short_industry = str_replace(short_industry, " and ", " & "))
         
top_10_industries <- top_10_industries %>% 
  mutate(Qindustry = factor(Qindustry, levels = Qindustry),
         position = cumsum(percent) - percent / 2,
         label = ifelse(percent >= 5, paste0(short_industry, "\n", round(percent), "%"), ""))


pie_chart(df                 = top_10_industries,
          column_name        = Qindustry,
          original_column_string = "Qindustry",
          text_position          = 100 - position,
          title_string           = "Respondent Industry",
          fill_legend_title      = "",
          colors                 = "Set3")

```

## Ethnicity

```{r Ethnicity, echo = TRUE}
# Multiple categories selected will be separated by commas, so we'll count those as multiple ethnicities
# This will probably miscategorize some open ends, but a quick analysis showed we have very few open ends
# with embedded commas.

ethnicities <-survey %>% group_by(Qethnicity_coded) %>% count(sort = TRUE)
num_ethnicity_responses <- sum(!is.na(survey$Qethnicity_coded))
ethnicities <- ethnicities %>% 
  mutate(percent = round(n / num_ethnicity_responses * 100, 0)) %>%
  ungroup() %>% 
  drop_na() %>% 
  arrange(desc(percent))

ethnicities <- ethnicities %>% 
 mutate(Qethnicity_coded = factor(as.character(Qethnicity_coded), levels = Qethnicity_coded),
         position = cumsum(percent) - percent / 2,
         label = ifelse(percent >= 2, paste0(Qethnicity_coded, "\n", round(percent), "%"), ""))

## Another customized pie chart here to use repelled labels to show small minority groups

ggplot(ethnicities) +
    geom_col(aes(x = 1, y = percent, fill = Qethnicity_coded), 
             width = 1, size = 1, color = "gray90", alpha = 0.4) +
    geom_text_repel(aes(x = 1.2, y = 100 - position, label = label),
              size = 3,
              hjust = 0.4,
              color = "black",
              direction = "y",
              force = 0.8
    ) +
    labs(x = NULL, y = NULL, 
         title = "Respondent Ethnicities",
         subtitle = question_text("Qethnicity"),
         caption = "2018 RStudio Learning R Survey",
         fill = "Selected Ethnicity") +
    guides(fill = guide_legend()) +
    scale_fill_brewer(palette = "Set1") +
#   scale_fill_viridis_d("Ethnicities", option = "A", direction = -1, alpha = 0.4) +
    theme(legend.key.width=unit(1, "cm")) +
    theme_classic() +
   coord_polar("y") +
#    coord_flip() +
    theme(axis.line = element_blank(),
         axis.text = element_blank(),
         axis.ticks = element_blank())

```

# Learning Questions

## What is Your Experience with R?

```{r R Experience so far, echo = TRUE}
# Multiple categories selected will be separated by commas, so we'll count those as multiple ethnicities
# This will probably miscategorize some open ends, but a quick analysis showed we have very few open ends
# with embedded commas.

r_experience <- tally_question(survey, Qr_experience, rm.na = TRUE, sort = FALSE)

r_experience <- r_experience %>% 
 mutate(Qr_experience_factor = factor(Qr_experience, levels = Qr_experience),
         position = cumsum(percent) - percent / 2,
         label = ifelse(percent >= 2, paste0(Qr_experience, "\n", round(percent), "%"), ""))

pie_chart(df                 = r_experience,
          column_name        = Qr_experience_factor,
          original_column_string = "Qr_experience",
          text_position          = 100 - position,
          title_string           = "R Experience",
          fill_legend_title      = "",
          colors                 = "Set1")

```

## How Did You Learn R?

```{r How Did You Learn R, echo = TRUE}
# We shorten down the given answers while categorizing open-text answers given for "Other"
# Just as we did for the gender and industry questions, we use small dictionary to do this
# However, because commas may appear in the open text respones, we'll use a tsv file instead of a csv one.

r_path_dictionary <- read_delim("data/r-path-dictionary.tsv", delim = "\t")
r_paths <- survey %>% select(Qr_learning_path) %>% left_join(r_path_dictionary, c("Qr_learning_path" = "Input"))
r_path_results <- tally_question(r_paths, Output, rm.na = TRUE, sort = TRUE)
answer_table <- r_path %>% select(Qr_learning_path, n)
# write_delim(answer_table, "data/r-path-dictionary.tsv", delim="\t")

r_path_results <- r_path_results %>% 
 mutate(Output = factor(Output, levels = Output),
         position = cumsum(percent) - percent / 2,
         label = ifelse(percent >= 2, paste0(Output, "\n", round(percent), "%"), ""))

pie_chart(df                 = r_path_results,
          column_name        = Output,
          original_column_string = "Qr_learning_path",
          text_position          = 100 - position,
          title_string           = "R Learning Path",
          fill_legend_title      = "",
          colors                 = "Set2")

```

## Respondent Ages By Learner Type

```{r R Learning Ages By Learner Type, echo = TRUE}

ggplot(filter(survey, !is.na(learner_type)), aes(x = 2018 - Qyear_born, group = learner_type)) +
  geom_density(aes(fill = learner_type), color="black", alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Respondent Ages (Density Plot)", 
         caption = "2018 RStudio Learning R Survey",
       x="Age", y="Density", fill = "Learner Type")
```

## Year Respondents Learned R

```{r Year respondents learned R, echo = TRUE}

ggplot(filter(survey, Qr_year >= 1990)) +
  geom_histogram(aes(Qr_year), binwidth = 1, fill = params$bar_colors, color="black") +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018)) +
  labs(title = "Year Started Learning R", 
       subtitle = question_text("Qr_year"),
       caption = "2018 RStudio Learning R Survey",
       x="Year", y="Count")
```


## Number of People You Work With That Know R By Learner Type

```{r Number R Team Members, echo = TRUE}
team_num_survey <- survey %>% filter(!is.na(learner_type))
ggplot(team_num_survey, aes(x = Qteam_r_users, group = learner_type)) +
  geom_density(aes(fill = learner_type), color="black", alpha = 0.5) +
  scale_x_continuous(trans = "log10") +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Number of Team Members Who Know R (Density Plot)", 
       subtitle = question_text("Qteam_r_users"),
       caption = "2018 RStudio Learning R Survey",
       x="Number", y="Density", fill = "Learner Type")
```

## Difficulty of Learning R based on Learner Type

```{r Ages By Learner Type, echo = TRUE}

ss <- survey %>% filter(!is.na(learner_type) & !is.na(Qr_difficulty_experienced))
ss_grouped <- ss %>% 
  group_by(learner_type, Qr_difficulty_experienced) %>% 
  count(Qr_difficulty_experienced)
ggplot(ss) +
  geom_bar(aes(x = Qr_difficulty_experienced, y = ..prop.., fill = learner_type, group = learner_type),
           position = position_dodge(width = 0.5),
           alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Difficulty in Learning R", 
       subtitle = question_text("Qr_difficulty_experienced"),
       caption = "2018 RStudio Learning R Survey",
 x="Respondent Rating (5 = Most Difficult)", y="", fill = "Learner Type")
```

## What is R Used For?

```{r R Used For, echo = TRUE}
# Breaks up uses into nested lists
uses <- survey %>% mutate(application = purrr::map(Qused_for, str_split, ", ")) %>% unnest()
tools_used <- uses %>% unnest() %>% group_by(application) %>% count(sort=TRUE)
# Count the non-NA entries now
use_responses <- survey %>% summarize(responses = sum(!is.na(Qused_for)))
tools_used <- tools_used %>% mutate(percent = round(n / use_responses$responses * 100)) %>% arrange(desc(percent))
app_plot <- head(tools_used, 15) %>% ungroup() %>% arrange(percent)

other <- tail(tools_used, -15) %>% 
  ungroup() %>% 
  summarize(application = "Other", 
            n = sum(n), 
            percent = round(n / use_responses$responses * 100))

app_plot2 <- rbind(other, app_plot) %>% drop_na()
app_plot2 <- app_plot2 %>% 
  mutate(application = factor(as.character(application), levels = application))

ggplot(app_plot2) +
  geom_col(aes(x=application, y = percent),
           fill =  params$bar_colors) +
  geom_text(aes(x=application, y = percent / 2, label = paste0(percent, "%")), 
            color="gray80", size=3, nudge_y = 1) +
  labs(title = "What Do You Use R For?", 
       subtitle = question_text("Qused_for"),
       caption = "2018 RStudio Learning R Survey",
x="Application", y="Percent") +
  coord_flip()

```

## Net Promoter Score

```{r Net Promoter Score, echo = TRUE}

recommend_totals <-survey %>% group_by(Qrecommend) %>% tally()
responses <- sum(recommend_totals$n, na.rm = TRUE)
recommend_totals <- recommend_totals %>% mutate(percent = n / responses * 100)
recommend_scores <- tibble(Qrecommend = 1:10) %>% left_join(recommend_totals, by = "Qrecommend")
nps <- round(sum(recommend_scores$percent[9:10], na.rm=TRUE) - 
               sum(recommend_scores$percent[1:6], na.rm=TRUE))  # Standard NPS formula (see Wikipedia)
npr <- nps %>% cut(breaks = c(-100, 0, 50, 75, 100), labels = c("Dangerous", "Good", "Excellent", "World Class"))

ggplot(recommend_scores) +
  geom_col(aes(x = Qrecommend, y = percent), fill = params$bar_colors, color="black") +
  geom_text(aes(x = Qrecommend, y = percent, label = paste0(round(percent), "%")), 
            color="black", size=3, nudge_y = 3) +
  scale_x_continuous(breaks = 1:10) +
    labs(title = question_text("Qrecommend"),
       subtitle = sprintf("Net Promoter Score = %s,        Rating: %s", nps, npr),
       caption = "2018 RStudio Learning R Survey",
       x="Score", y="Count")
```

## What Tools Do You Use With R?

```{r R Tools Used, echo = TRUE}

uses <- survey %>% transmute(tools = str_replace_all(Qtools_with_r, "i.e., ", ""))
num_tool_responses <- sum(!is.na(survey$Qtools_with_r))
tools_used <- split_and_aggregate(uses, tools)
tools_used <- tools_used %>% 
  select(-num_responses) %>% 
  mutate(percent = round(n / num_tool_responses * 100))
top_15_items <- top_n_choices(tools_used, items, num_tool_responses, 15) %>% arrange(percent)
top_15_items <- top_15_items %>%
  mutate(items = factor(as.character(items), levels = items))

# Plot
ggplot(top_15_items) +
  geom_col(aes(x=items, y = percent),
           fill =  params$bar_colors) +
  geom_text(aes(x=items, y = percent , label = paste0(percent, "%")), 
            color="gray30", size=3, nudge_y = 3) +
  labs(title = "What Tools Do You Use With R?", 
       subtitle = question_text("Qtools_with_r", 55),
       caption = "2018 RStudio Learning R Survey",
       x="", y="") +
  coord_flip() +
  theme(axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank())

```

