---
title: "Learning R Survey Results"
author: "Carl Howe, RStudio"
date: "14 December 2018"
output:
  beamer_presentation: default
  ioslides_presentation:
    css: styles.css
    logo: rstudio.png
    smaller: yes
    widescreen: yes
subtitle: 14 December 2018
params:
  bar_colors: dark green
  language: English
---

# Learning R Survey Results

## Methodology

- Survey conducted between December 6 and December 31, 2018
- Survey was fielded in both English and Spanish versions
- Respondents solicited from
    + community.rstudio.com
    + Twitter followers of RStudio employees
    + reddit.com/datascience
- Survey results are not representative of any broader population
- Data and processing scripts can be found the repo at the end of this talk
- Data and scripts are open source

## Setup

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(googlesheets)
library(tidyverse)
library(RColorBrewer)
library(gendercodeR)

# First, we'll just read in the data and fix all the question names, which in Google Forms, are literally
# the full text of the question. I've put short question names on another sheet of the results Google Sheet, so
# we do the substitution here.

mysheets <- gs_ls() # will authenticate
if (params$language == "English") {
  survey_handle <- gs_title("Learning R Internet Survey")
} else {
  survey_handle <- gs_title("Learning R Internet Survey (Spanish Version)")
}
 # get the survey results
survey <- survey_handle %>%
  gs_read(ws = "Form Responses 1")

## Unfortunately, this read brings in the question text as all the column names. That's not good.
## We'll save them in question_text and put shorter, more friendly names in instead.

survey_questions <- survey_handle %>% gs_read(ws="Question Names")
names(survey) <- survey_questions$Question_name
respondents <- nrow(survey)
survey_save <- survey                # in case I mess up survey, I can check it against this
est <- locale(tz="US/Eastern")
survey$Qtime <- parse_datetime(survey$Qtime, format = "%m/%d/%Y %H:%M:%S", locale=est)

# This learner_type variable defines if someone is a recent R learner or not for faceting
survey <- survey %>% 
  mutate(learner_type = ifelse(Qr_year < 1900, NA, ifelse(Qr_year <= 2016, "Early Learner", "Recent Learner")))

# This codes the open-text gender question into a more friendly form for analysis
survey$Qgender <- tolower(survey$Qgender)
survey_gendercoded <- genderRecode(survey, method = "narrow", genderColName = "Qgender", outputColName = "Qgender_coded")

theme_set(theme_bw())

```

## Responses over time

```{r responses over time}
survey <- survey %>% mutate(number_responses = cumsum(!is.na(Qtime)))
ggplot(survey) + 
  geom_line(aes(x=Qtime, y=number_responses)) + 
  ggtitle(sprintf("%s Survey Responses Over Time (N = %d)",params$language, respondents)) + 
  labs(y = "Time", x = "Response Date")

```


## Respondent Countries

```{r Respondent Countries, echo = TRUE}
countries <- survey %>% count(Qcountry, sort=TRUE) %>% drop_na()
names(countries) <- c("region", "value")
countries <- countries %>% mutate(region = tolower(region))
unknown_countries <- c("singapore", "korea", "south", "serbia and montenegro", "hong kong", "andorra", "bahrain", "christmas island", "macau", "mauritius", "puerto rico", "tanzania")
choro_countries <- countries %>% filter(!(region %in% unknown_countries))
choro_countries <- choro_countries %>% mutate(region = str_replace_all(region, "korea, south", "south korea"))

library(choroplethr)
library(choroplethrMaps)
data(country.map)

# country_choropleth(choro_countries, title = "", legend = "", num_colors = 8,
#  zoom = NULL)

plot_countries <- country.map %>% left_join(choro_countries, by = "region")
ggplot(data = plot_countries, aes(long, lat, group=group)) + 
  geom_polygon(aes(fill = value), color = "gray", size=0.1) +
  scale_fill_gradient(trans = "log10", low = "dark blue", high = "yellow") +
  coord_fixed(1.3) +
  scale_x_continuous(breaks = seq(-150, 150, 60)) +
  scale_y_continuous(breaks = seq(-60, 90, 30), limits = c(-60, 90)) +
  labs(title = "Geographical Distribution of Survey Respondents", x = "", y = "", fill = "Responses")
```
## Respondent Ages

```{r Respondent Ages, echo = TRUE}

ggplot(survey) +
  geom_histogram(aes(2018 - Qyear_born), binwidth = 5,  fill = params$bar_colors, color="black") +
  scale_x_continuous(breaks = seq(15,80,5)) +
  labs(title = "Respondent Ages (Histogram)", x="Age", y="Count")
```

## Respondent Ages By Learner Type

```{r R Learning Ages By Learner Type, echo = TRUE}

ggplot(filter(survey, !is.na(learner_type)), aes(x = 2018 - Qyear_born, group = learner_type)) +
  geom_density(aes(fill = learner_type), color="black", alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Respondent Ages (Density Plot)", x="Age", y="Density", fill = "Learner Type")
```

## Year Respondents Learned R

```{r Year respondents learned R, echo = TRUE}

ggplot(filter(survey, Qr_year >= 1990)) +
  geom_histogram(aes(Qr_year), binwidth = 1, fill = params$bar_colors, color="black") +
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2018)) +
  labs(title = "Year Started Learning R (Histogram)", x="Year", y="Count")
```


## Number of People You Work With That Know R By Learner Type

```{r Number R Team Members, echo = TRUE}
team_num_survey <- survey %>% filter(!is.na(learner_type))
ggplot(team_num_survey, aes(x = Qteam_r_users, group = learner_type)) +
  geom_density(aes(fill = learner_type), color="black", alpha = 0.5) +
  scale_x_continuous(trans = "log10") +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Number of Team Members Who Know R (Density Plot)", x="Number", y="Density", fill = "Learner Type")
```

## Difficulty of Learning R based on Learner Type

```{r Ages By Learner Type, echo = TRUE}

ss <- survey %>% filter(!is.na(learner_type) & !is.na(Qr_difficulty_experienced))
ss_grouped <- ss %>% 
  group_by(learner_type, Qr_difficulty_experienced) %>% 
  count(Qr_difficulty_experienced)
ggplot(ss) +
  geom_bar(aes(x = Qr_difficulty_experienced, y = ..prop.., fill = learner_type, group = learner_type),
           position = position_dodge(width = 0.5),
           alpha = 0.5) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Difficulty in Learning R", x="Respondent Rating (5 = Most Difficult)", y="", fill = "Learner Type")
```


## Respondent Gender

```{r Respondent Gender, echo = TRUE}
# survey_gendercoded <- genderRecode(survey, genderColName = "Qgender", outputColName = "Qgender_coded")
# genders <- survey_gendercoded %>% count(Qgender_coded, sort = TRUE)
## Much work to do here
```

## What is R Used For?

```{r R Used For, echo = TRUE}
# Breaks up uses into nested lists
uses <- survey %>% mutate(application = map(Qused_for, str_split, ", ")) %>% unnest()
tools_used <- uses %>% unnest() %>% group_by(application) %>% count(sort=TRUE)
# Count the non-NA entries now
use_responses <- survey %>% summarize(responses = sum(!is.na(Qused_for)))
tools_used <- tools_used %>% mutate(pct = round(n / use_responses$responses * 100)) %>% arrange(desc(pct))
app_plot <- head(tools_used, 15) %>% ungroup()
other <- tail(tools_used, -15) %>% 
  ungroup() %>% 
  summarize(application = "Other", 
            n = sum(n), 
            pct = round(n / use_responses$responses * 100))
app_plot <- rbind(app_plot, other) %>% drop_na()
ggplot(app_plot) +
  geom_col(aes(x=reorder(application, pct), y = pct),
           fill =  params$bar_colors) +
  geom_text(aes(x=reorder(application, pct), y = pct / 2, label = paste0(pct, "%")), 
            color="gray80", size=3) +
  labs(title = "What Do You Use R For?", x="Application", y="Percent") +
  coord_flip()

```
## Net Promoter Score

```{r Net Promoter Score, echo = TRUE}
recommend_totals <-survey %>% group_by(Qrecommend) %>% tally()
responses <- sum(recommend_totals$n, na.rm = TRUE)
recommend_totals <- recommend_totals %>% mutate(percent = n / responses * 100)
recommend_scores <- tibble(Qrecommend = 1:10) %>% left_join(recommend_totals, by = "Qrecommend")
nps <- round(sum(recommend_scores$percent[9:10], na.rm=TRUE) - 
               sum(recommend_scores$percent[1:6], na.rm=TRUE))  # Standard NPS formula (see Wikipedia)
npr <- nps %>% cut(breaks = c(-100, 0, 50, 75, 100), labels = c("Dangerous", "Good", "Excellent", "World Class"))


ggplot(survey) +
  geom_histogram(aes(x = Qrecommend), binwidth = 1,  fill = params$bar_colors, color="black") +
  scale_x_continuous(breaks = 1:10) +
    labs(title = sprintf("How Likely Are You To Recommend R To A Friend Or Colleague?\nNet Promoter Score = %s,        Rating: %s", nps, npr), x="Score", y="Count")
```
## What Tools Do You Use With R?

```{r R Tools Used, echo = TRUE}
# Breaks up uses into nested lists, split by commas
uses <- survey %>% mutate(application = map(Qtools_with_r, str_split, ", ")) %>% unnest()
tools_used <- uses %>% unnest() %>% group_by(application) %>% count(sort=TRUE)

# Count the non-NA entries now
use_responses <- survey %>% summarize(responses = sum(!is.na(Qused_for)))
tools_used <- tools_used %>% mutate(pct = round(n / use_responses$responses * 100)) %>% arrange(desc(pct))

# Now take these responses and only show the top N, aggregating the rest into an Other category
# This will be a function soon.

app_plot <- head(tools_used, 15) %>% ungroup()
other <- tail(tools_used, -15) %>% 
  ungroup() %>% 
  summarize(application = "Other", 
            n = sum(n), 
            pct = round(n / use_responses$responses * 100))
app_plot <- rbind(app_plot, other) %>% drop_na() 

# Plot
ggplot(app_plot) +
  geom_col(aes(x=reorder(application, pct), y = pct),
           fill =  params$bar_colors) +
  geom_text(aes(x=reorder(application, pct), y = pct / 2, label = paste0(pct, "%")), 
            color="gray80", size=3) +
  labs(title = "What Tools Do You Use With R?", x="Tool", y="Percent") +
  coord_flip()

```


