---
title: Using a custom dictionary
author: "Emily Kothe"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This gendercodeR package allows rapid recoding of freetext gender responses 
using an in-built dictionary. Individuals using the package may wish to extend 
or overwrite elements of the dictionary where they disagree with the coding 
decisions made by the package authors and/or wish to include recoding options 
that are not included in the in-built dictionary. A custom dictionary can be 
created by the user and provided to the `genderRecode()` function for this purpose. 

## Adding new items to the dictionary

For example, when collecting data from Spanish speaking participants a 
participant who enters their gender as "masculino" will not be recoded using 
the in-built dictionary.

```{r}

df <- data.frame(stringsAsFactors=FALSE,
      gender = c("male", "MALE", "mle", "mail", "masculino", "female", "enby"),
         age = c(34L, 37L, 77L, 52L, 68L, 67L, 83L)
)

df

library(gendercodeR)

genderRecode(df)
```

The custom dictionary should be a data frame with two columns. Column 1 should 
include the input text (i.e. the response provided by the participant) and 
Column 2 should include the desired output text (i.e. the gender category the 
input should be converted to).

```{r}

custom_dictionary <- data.frame(stringsAsFactors=FALSE,
       Input = c("masculino", "femenina"),
      Output = c("male", "female")
)

custom_dictionary
```

The custom dictionary is appended to the in-built dictionary and participants 
are recoded as desired.

```{r}
genderRecode(df, customDictionary = custom_dictionary)
```

## Overwriting existing entries in the dictionary

The narrow recoding method recodes participants who indicate that they are 
intersex are recoded to "Sex and Gender Diverse". In some contexts researchers 
may wish to use narrow recoding but overwrite this particular behaviour so that 
individuals who report they are coded as intersex but the other entries in the 
dictionary are preserved.^[The package authors wish to acknowledge that the 
narrow categories make opionated assumptions and that the broad recoding will 
generally be more appropriate.]

```{r}

df <- data.frame(stringsAsFactors=FALSE,
      gender = c("male", "MALE", "mle", "mail", "intersex", "female", "enby"),
         age = c(34L, 37L, 77L, 52L, 68L, 67L, 83L)
)

df

genderRecode(df, method = "narrow")
```

A custom defintion can be provided and then be applied.

```{r}

custom_dictionary <- data.frame(stringsAsFactors=FALSE,
       Input = c("intersex"),
      Output = c("intersex")
)

genderRecode(df, method = "narrow", customDictionary = custom_dictionary)
```
